### 2024-03-25 paper/notes_for_intro.tex

General notes:

RG 2019 looked at the simulated morphology of illustristng galaxies and found that for M*<1e9.5, the morphologies are "less reliable" -> motivates using pair fractions for merger rate studies

Further work that would be nice to do would be to see if all the timescale stuff stays the same for minor mergers, and for smaller bins from 1-1:4

O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate
Jiang 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq. 4\&5). Also quantified the circularity of orbits as a function of mass ratio.
Jiang 2014

Paper:
Author:
What's the connection:
Other notes:

PRIORITY: 
    Author: O'Leary 2021
    Paper: Emerge: Constraining merging probabilities and timescales of close galaxy pairs 
    What's the connection:
    Other notes: 
    Empirical models (EMERGE) to probe the dependencies of pair merging probabilities and merging timescales. Also show that some commonly used pair selection criteria may not represent a suitable sample of galaxies to reproduce underlying merger rates. Seem to talk about both observation timescales and merger timescales, and I'm not sure which one they focus on.
    Quantified probability of a "close pair" merging before z=0 and dependence on rsep-proj, LOS vel, and z. 
    On what timescale will an observed pair merge, and for how long is that pair observable given the selection criteria. 
    What determines pair observation timescales (might be a followup from their 2020 paper) 
    Notes that papers that attempt to come up with fitting functions for timescales etc. will all be wrong for pairs w different selection criteria.
    -> Maybe motivates that we should say that future extensions of our work may lead to the development of fitting functions that can be applied to any system? (is this true?)

    Author: Ventou2017
    Paper: 
    What's the connection: 
    Other notes:

    Author: Patton2024
    Paper: 
    What's the connection:
    Other notes:

    Author: Guzman-Ortega 2022
    Paper: Morphological signatures of mergers in the TNG50 simulation and the Kilo-Degree Survey: the merger fraction from dwarfs to Milky Way-like galaxies
    What's the connection:
    Other notes:

OTHER:
    Author: Wang2020 (+Pearson, Rodriguez-Gomez)
    Paper: Towards a consistent framework of comparing galaxy mergers in observations and simulations
    What's the connection: Looked at the major merger fraction of stuff at low-z. Their stellar mass criteria did not cover our range of interest for low-mass systems
    Other notes: Mass dependence of the galaxy major merger fraction at low z. Make synthetic training images from TNG and apply to KiDS images from GAMA. Find that major merger fraction decreases slightly with increasing stellar mass (1e9.5-1e11.5) -> 
    major mergers defined as happening in the next 1Gyr or in the previous 500Myr.

    Paper: 
    Author: RG2017
    What's the connection: 
    Other notes:

    Paper: 
    Author: Patton2020
    What's the connection:
    Other notes:

    Paper: 
    Author: Patton2000
    What's the connection:
    Other notes:

    Paper: Emerge: Empirical predictions of galaxy merger rates since $z \sim 6$ 
    Author: O'Leary 2020
    What's the connection:
    Other notes:

    Paper: SATELLITES IN MILKY-WAY-LIKE HOSTS: ENVIRONMENT DEPENDENCE AND CLOSE PAIRS
    Author: González2013
    What's the connection:
    Other notes:

    Paper: Observational Constraints on the Merger History of Galaxies since $z\sim6$: Probabilistic Galaxy Pair Counts in the CANDELS Fields
    Author: Duncan 2019
    What's the connection:
    Other notes:

    Paper: THE MILLENNIUM GALAXY CATALOGUE: THE CONNECTION BETWEEN CLOSE PAIRS AND ASYMMETRY; IMPLICATIONS FOR THE GALAXY MERGER RATE
    Author: De Propris 2007
    What's the connection:
    Other notes:

    Paper: Galaxy interactions in IllustrisTNG-100, I: The power and limitations of visual identification
    Author: Blumenthal 2020
    What's the connection:
    Other notes:

    Paper: Small-Scale Challenges to the ΛCDM Paradigm
    Author: Bullock 2017
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes: